---
title: "choice model data preparation"
author: "Tobias Schwoerer"
date: "March 22, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

Importing data, setting aside individual specific characteristics, identifying each pilot's stated 2016 priority region based on the maximum proportion of flights the pilot took into that region
```{r}
library(tidyr)
library(dplyr)
library(data.table)

cdata <- read.csv("Data/new_data_w_elodeaalt3xprivate110718.csv")

#individual pilot specific data 
indvData <- cdata%>%
  subset(set==1)%>%
  select(id, alt, cost, age, weight, flights, income, flightsnew, sharenew, choice)

#revealed priority region (from set 1, recorded, nonhypothetical)
regionPref <- indvData%>%
  group_by(id)%>%
  arrange(desc(sharenew),alt) %>%
  slice(1)%>%
  select(id,alt)
```


pilots who have more than one region selected in set 2
```{r}
multi <- cdata%>%
  subset(set==2)%>%
  group_by(id)%>%
  summarise(sum=sum(choice))%>%
  subset(sum>1)
#adding priority region 
multi <- multi%>%
  left_join(regionPref, by="id")
```

#Creating truly binary choice data set
here the region "chosen" in the second hypothetical (contingent) part of the panel is based on the maximum proportion of flights into the region. 
```{r}
#Setting choice in set 1 equal to priority region from above, creating new variable choice2 to do this
cdata2 <- cdata%>%
  left_join(regionPref, by="id")

cdata2$choice2 <- with(cdata2, ifelse(set==1 & alt.x==alt.y, 1,0)) 
```

Note, choice in set 2 is already set to the maximum proportion of flights into that region, except where there are equal proportions.
These pilots are contained in multi above. For them, the priority region from above was inferred.
```{r}
cdata2 <- cdata2%>%
  left_join(multi)

#setting choice in set 2, creating new variable choice3 to do this
cdata2$choice3 <- with(cdata2, ifelse(set==2 & alt.x==alt, 1,0)) 

#combining choice, choice2, and choice3 by creating final choice variable choiceF
cdata2$choiceF <- with(cdata2, ifelse(is.na(choice3),choice,ifelse((set==1), choice2,choice3)))
```

testing if calculations are correct
```{r}
test <- cdata2%>%
  group_by(id,set)%>%
  summarise(sum(choiceF))
```
Correct, there is only one region selected per set now

#Creating dataset for estimation
```{r}
#eliminating columns we won't need
cdata3 <- cdata2%>%
  select(-c("choice","choice2","choice3","alt","sum","alt.y","wgt1","wave","contflights","contflightsw","wgtc","elightsw","wgte","share","weight","sharenew","flights","flightsw","eflights"))

#renaming columns
setnames(cdata3, old=c("alt.x","elodeanew","flightsnew","flightsneww","choiceF"), new=c("alt", "elodea","flights","flights_w","choice"))

#writing file to dropbox
write.csv(cdata3, file="D:/Dropbox/DATA/2015_Schwoerer_floatplane_survey/choice_data032219.csv")
```