---
title: "AIS seaplane transmission risk "
author: "Tobias Schwoerer"
date: "November 20 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Hierarchical model using rstanarm for estimating destination specific probability distribution of introducing AIS to lake j  
With help from Eric Ward, NOAA
## Useful references: Gelman et al. 2013. Bayesian Data Analysis 3rd ed. , Carpenter et al. 2016. Stan: A probabilistic programming language. Journal of Statistical Software

#1 Set up
load packages, read in and prepare data for analysis 
```{r}
library(shinystan)
library(bayesplot)
library(ggplot2)
library(cowplot)
library(rstan)
library(dplyr)
library(tidyr)
library(foreign)
library(data.table)
library(tidyverse)
library(loo)

theme_set(theme_classic(base_size = 9))
#importing data archived at https://doi.org/10.18739/A25H7BV1C 
#data <- read.csv(url("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3Aaac2cc93-04f3-4e5c-b36b-01120993f1d8",      method="libcurl"))
#ref  <-  read.csv(url("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3A8ed27563-55f2-4273-a272-09406d85419b",method="libcurl")) 
lakeList <- read.csv("D:/Dropbox/DATA/2015_Schwoerer_floatplane_survey/FlightPattern/LakeList.csv")
flights <- read.csv("D:/Dropbox/DATA/2015_Schwoerer_floatplane_survey/FlightPattern/flight_pattern_long.csv")
#lakeList <- read.csv("C:/Users/Toby/Dropbox/DATA/2015_Schwoerer_floatplane_survey/FlightPattern/LakeList.csv",stringsAsFactor=F)
#flights <- read.csv("C:/Users/Toby/Dropbox/DATA/2015_Schwoerer_floatplane_survey/FlightPattern/flight_pattern_long.csv",stringsAsFactor=F)
```

#setting a common lakeID for all lakes
```{r}
#eliminating non-freshwater take off indicated by Type==3 in LakeList
s_list <- read.csv("D:/Dropbox/DATA/2015_Schwoerer_floatplane_survey/FlightPattern/Start_ID_Fetch_data.csv",stringsAsFactor=F)
lakeList2 <- lakeList%>%
  filter(Type!=3)%>%
  left_join(s_list, by=c("SeaplaneBase"="StartID"))
#recoding Stlk_# with Lk_#
lakeList2$LakeID  <- with(lakeList2, ifelse( str_detect(LakeID,"Stlk_"), UniqueID,LakeID))
lakeList2 <- lakeList2[,1:8] 
write.csv(lakeList2, file="D:/Dropbox/DATA/2015_Schwoerer_floatplane_survey/FlightPattern/lakeList2.csv")
```

#Data preparation
```{r}
#eliminate flights from runways (amphibious planes)
runway <- lakeList%>%
  filter(Type==3)
v <-runway$LakeID
#drop rows associated with runway start location
flights <- flights[!(flights$StartID %in% v), ]

#create ElodeaO and ElodeaD indicating whether origin or destination is known elodea source
starts <- filter(lakeList, lakeList$SeaplaneBase != "")
starts <- starts[,c("Elodea","SeaplaneBase")]
flights <- flights%>%
  left_join(starts,by=c("StartID"="SeaplaneBase"))
names(flights)[names(flights)=="Elodea"] <- "ElodeaO"
dest <- lakeList[,c("Elodea","LakeID")]
flights <- flights%>%
  inner_join(dest,by=c("DestID"="LakeID"))
names(flights)[names(flights)=="Elodea"] <- "ElodeaD"

flights$Oeflights <- with(flights,ifelse(ElodeaO==0&ElodeaD==1|ElodeaO==1&ElodeaD==1,AnnualFl_1,0))
flights$Deflights <- with(flights,ifelse(ElodeaO==1&ElodeaD==0|ElodeaO==1&ElodeaD==1,AnnualFl_1,0))

#summarizing origin 
origin <- flights%>%
  group_by(StartID)%>%
  summarise(eflights=sum(Oeflights),
            flights=sum(AnnualFl_1))
ids <- lakeList2[,c("SeaplaneBase","LakeID")]
origin <- origin%>%  
  left_join(ids, by=c("StartID"="SeaplaneBase"))
#drop StartID column
origin <- origin%>%
  select(-StartID)

#summarizing destinations
destinations <- flights%>%
    group_by(DestID)%>%
    summarise(eflights=sum(Deflights),
            flights=sum(AnnualFl_1))
names(destinations)[names(destinations)=="DestID"] <- "LakeID"

#combining
data <- bind_rows(origin,destinations)%>%
  group_by(LakeID)%>%
    summarise(eflights=sum(eflights),
            flights=sum(flights))

J <- nrow(data)
n <- data$flights
y <- data$eflights
```

Testing for heterogeneity of proportions
```{r}
library(rstatix)
# Homogeneity of proportions between groups
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
# H0: the proportion of eflights is similar for each lake
xtab <- as.table(rbind(y, n-y))

# Compare the proportions of eflights between lakes
propTest <- prop_test(xtab, detailed=TRUE)

# Pairwise comparison between lakes
pwPropTest <- pairwise_prop_test(xtab, p.adjust.method = "hochberg")
```


#2 Partial pooling Model
Following: https://mc-stan.org/users/documentation/case-studies/pool-binary-trials-rstanarm.html#observed-vs.estimated-chance-of-success
```{r}
library(rstanarm)
Lake <- data$LakeID
lake_avgs <- y / n
tot_avg <- sum(y) / sum(n)
intro_avg <- function(x) print(format(round(x, digits = 3), nsmall = 3), quote = FALSE)
summary_stats <- function(posterior) {
  x <- invlogit(posterior)  # log-odds -> probabilities
  t(apply(x, 2, quantile, probs = c(0.1, 0.5, 0.9))) 
}
options(mc.cores = parallel::detectCores())
SEED <- 101
wi_prior <- normal(0, 1.4)  # this flat prior from King et al. (2009), also used weakly informative prior on log-odds normal(-1, 1), which is slightly informative on smaller values but compared to this flat prior was not as good in the model comparison below, see hist(plogis(rnorm(1000,-1,1)),100)

fit_partialpool <- 
  stan_glmer(cbind(y, n - y) ~ (1 | Lake), data = data,  family = binomial("logit"),
             prior_intercept = wi_prior, seed = SEED, adapt_delta = 0.99, iter=10000) #The left-hand side of the formula specifies the binomial outcome by providing the number of successes (hits) and failures (flights from non-sources) for each lake, and the right-hand side indicates that we want an intercept-only model.
```

#3 Plotting results table
TASK turn this into a dataframe and combine with other ref info, keep model evaluation parameters Rhat etc.
TASK 2: color know elodea locations
```{r}
# shift each Lake's estimate by intercept (and then drop intercept)
shift_draws <- function(draws) {
  sweep(draws[, -1], MARGIN = 1, STATS = draws[, 1], FUN = "+")
}
alphas <- shift_draws(as.matrix(fit_partialpool))
partialpool <- summary_stats(alphas)
partialpool <- partialpool[-nrow(partialpool),]
rownames(partialpool) <- as.character(data$LakeID)
intro_avg(partialpool)

#create dataframe and moving rownames into columns
partialpoolDF <- as.data.frame(partialpool)
partialpoolDF <- rownames_to_column(partialpoolDF, var="LakeID")

#save summary of fitted model as data.frame
summary(fit_partialpool)
diag_partial <-as.data.frame(summary(fit_partialpool))
#moving rownames into column
diag_partial <- rownames_to_column(diag_partial, var="LakeID")
#extracting intercept, sigma, mean PPD, log-posterior to save for later
drop_rows <- diag_partial[(diag_partial$LakeID %in% c("(Intercept)","Sigma[Player:(Intercept),(Intercept)]","mean_PPD","log-posterior")), ] 
#dropping the above
diag_partial <- diag_partial[ !(diag_partial$LakeID %in% c("(Intercept)","Sigma[Player:(Intercept),(Intercept)]","mean_PPD","log-posterior")), ] #cleaning up LakeID by first extracting string that starts with Lk and ends with ]
diag_partial$LakeID <- str_extract_all(diag_partial$LakeID, "LK.+]")
diag_partial$LakeID <- substr(diag_partial$LakeID, 1, nchar(diag_partial$LakeID)-1)
#drop columns showing log-odds
diag_partial <- diag_partial %>%
  select(LakeID, mcse, n_eff, Rhat)

#joining dataframe to create results table incl. diagnostics, except sigma, mean PPD, log-posterior
partialpoolResults <- diag_partial%>%
  left_join(partialpoolDF, by="LakeID")
```


#Complete pooling model
```{r}
fit_pool <- 
  stan_glm(cbind(y, n - y) ~ 1, data = data, family = binomial("logit"),
             prior_intercept = wi_prior, seed = SEED, adapt_delta = 0.99, iter=2000)

invlogit <- plogis  # function(x) 1/(1 + exp(-x))
summary_stats <- function(posterior) {
  x <- invlogit(posterior)  # log-odds -> probabilities
  t(apply(x, 2, quantile, probs = c(0.1, 0.5, 0.9))) 
}

pool <- summary_stats(as.matrix(fit_pool))  # as.matrix extracts the posterior draws
pool <- matrix(pool,  # replicate to give each player the same estimates
               nrow(data), ncol(pool), byrow = TRUE, 
               dimnames = list(data$LakeID, c("10%", "50%", "90%")))
intro_avg(pool)
```

#no pooling model
```{r}
fit_nopool <- update(fit_pool, formula = . ~ 0 + Lake, prior = wi_prior)
nopool <- summary_stats(as.matrix(fit_nopool))
rownames(nopool) <- as.character(data$LakeID)
intro_avg(nopool)

```

#Plotting predicted against the data, Posterior Medians and 80% Intervals
```{r}
library(ggplot2)
models <- c("no pooling","partial pooling")  #"complete pooling"
estimates <- rbind(nopool,partialpool)  #pool, nopool, 
colnames(estimates) <- c("lb", "median", "ub")
plotdata <- data.frame(estimates, 
                       observed = rep(lake_avgs, times = length(models)), 
                       model = rep(models, each = J), 
                       row.names = NULL)

ggplot(plotdata, aes(x = observed, y = median, ymin = lb, ymax = ub)) +
  geom_hline(yintercept = tot_avg, color = "lightpink", size = 0.75) +
  geom_abline(intercept = 0, slope = 1, color = "skyblue") + 
  geom_linerange(color = "gray60", size = 0.75) + 
  geom_point(size = 2.5, shape = 21, fill = "gray30", color = "white", stroke = 0.2) + 
  facet_grid(. ~ model) +
  coord_fixed() +
  scale_x_continuous(breaks = c(0.25,0.50,0.75,1.0)) +
  labs(x = "Observed y / n", y = "Predicted chance of AIS introduction") 

ggsave("STAN/flights_model/figures/post_med_compare.bmp", plot = last_plot(), device = "tiff", path = NULL,
  scale = 1, width = 140, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)
```

#Cross-validation for model checking and comparison
approximating the expected log predictive density (the lower the better) in the third column of the output. Specifically, it is the leave-one-out (loo) approximation to the log predictive density. Even though this approximation is only asymptotically valid, as it likely underestimates the expected log predictive density, the relative ranking of the models is the same as if it would be correctly calculated. https://mc-stan.org/users/documentation/case-studies/pool-binary-trials-rstanarm.html#partial-pooling
```{r}

loo_compare(loo(fit_partialpool), loo(fit_partialpool_wi))

loo_compare(loo(fit_partialpool), loo(fit_pool), loo(fit_nopool))
```
               elpd_diff se_diff 
fit_partialpool      0.0       0.0
fit_nopool       -1382.5      40.9
fit_pool        -11271.8    1835.2

The partial pool model with the flat non-informative prior outperforms the model with the weakly informative prior. 
                   elpd_diff se_diff
fit_partialpool     0.0       0.0   
fit_partialpool_wi -7.7       2.6



#Predicting new observations



#3 Model output
```{r}  
#Save posterior summary statistics (contains sampling diagnostics)
fit_summary <- summary(datafit)
fit_summary.df <- as.data.frame(fit_summary$summary)
fitS <- setDT(fit_summary.df, keep.rownames = "X")
fitS <- fitS[1:727,]
results <- fitS%>%
  left_join(ref, by=c("X"="ï..modelID"))

#Sort by median predicted introduction rate and inverse variance
results <- results[order(-results$X50.,results$sd),]
results <- results%>% 
  mutate(rank = row_number())

#Move columns for easier reading starting with Rank, UniqueID, 
results <-results[c("rank","UniqueID","DestName","Name","RegionName","DestType","Lat", "Long","HUC8","GMU","mean","se_mean","sd","X2.5.","X25.","X50.","X75.","X97.5.","n_eff","Rhat","WithinNPS","WithinRefu","Suitab","Fetch_m","ElevMEAN","flights","eflights","X")]
write.csv(results, file="/STANresults.csv",row.names = FALSE)
```


#4 Model diagnostic and validation
##4.1 Posterior predictive check
plotting observed introduction rates versus posterior medians with 95% posterior intervals
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
sdata <- read.csv("STAN/flights_model/fit_summary3.csv")
data <-  read.csv("STAN/floatplanedata4.csv")
data$rate <- data$eflights/data$flights

theme_set(theme_classic(base_size = 7))

##renaming columns of above data frame
names(sdata)[5:9] <- c("perc3","perc25","perc50","perc75","perc97")
#selecting first 729 rows for the draws
sdata2 <- sdata[1:727,]
sdata2 <- cbind(sdata2,data)

ScatterFig <-  ggplot(sdata2, aes(x=sdata2$rate,y=sdata2$perc50)) +
  geom_pointrange(aes(ymin=sdata2$perc3, ymax=sdata2$perc97)) + 
  xlab("observed introduction rate")+ ylab("95% posterior interval for theta(j)") +
  ylim(0,1)+ xlim(0,1) + geom_jitter(height=0,width=0.1)+ geom_abline(intercept = 0, slope = 1,colour='red')+
  theme(axis.title.x = element_text(face="bold", colour="#000000", size=7),
        axis.text.x  = element_text(face="bold", angle=0, vjust=0.5, size=7)) +
  theme(axis.title.y = element_text(face="bold", colour="#000000", size=7),
        axis.text.y  = element_text(face="bold", angle=360, vjust=0.5, size=7))+
  theme(panel.background = element_rect(fill = 'white', colour="black"))
ScatterFig
ggsave("STAN/flights_model/figures/datafit3Figures/ScatterFig.bmp", plot = last_plot(), device = "bmp", path = NULL,
  scale = 1, width = 90, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)
```

##4.2 Model checking
external validation
```{r}
data_sims <- extract(datafit3, permuted=TRUE)
theta_rep <- array(NA,c(n_sims,J))
y_rep <- array(NA,c(n_sims,J))
for (s in 1:n_sims){
  theta_rep[s,] <- rbeta(J,data_sims$alpha[s],data_sims$beta[s])
  y_rep[s,] <- rbinom(J,n_sims,theta_rep[s,])
}

 #displaying replicated data
par(mfrow=c(5,4),mar=c(4,4,2,2))
hist(y/n,xlab="",main="observed")
for (s in 1:19)
  hist(theta_rep[s,], xlab="",main=paste("theta_rep",s))
 #Saved figure manually for Supplementary File

 #calculating mean and SD for replicated new data
theta_rep_mean <- mean(theta_rep)
theta_rep_mean
theta_rep_sd <- sd(theta_rep)
theta_rep_sd

ratios_cp <- neff_ratio(datafit3)
mcmc_neff(ratios_cp, size = 2)
```
Results:
theta_rep_mean = 0.236
theta_rep_sd = 0.361

Also, the ratios are very low, below 0.1 which may hint torwards the need to reparameterize. See: https://cran.r-project.org/web/packages/bayesplot/vignettes/visual-mcmc-diagnostics.html#rhat-potential-scale-reduction-statistic

##4.3 Convergence check
Rhat, traceplot, energy plot, pairs plot 
```{r}
#number of lakes with Rhat<=1.05 Rhat<=1.1 and Rhat>1.1
df <- fit_summary.df[1:727,]#subsetting to just include the thetas for 727 lakes
df <- cbind(df,ref)
nrow(subset(df,Rhat<=1.05))
nrow(subset(df,Rhat<=1.1))
nrow(subset(df,Rhat>1.1))


#traceplots
traceplot <- stan_trace(datafit3, pars = c("theta[322]","theta[324]","theta[518]","avg"), inc_warmup = TRUE, nrow = 9)
levels(traceplot$data$parameter)[1] <- "Martin Lake"
levels(traceplot$data$parameter)[2] <- "McKinley Lake"
levels(traceplot$data$parameter)[3] <- "Alexander Lake"
levels(traceplot$data$parameter)[4] <- "Average"
traceplot <- traceplot + xlab("iteration") + ylab("posterior for theta(j)")
traceplot

ggsave("STAN/flights_model/figures/datafit3Figures/traceplot.bmp", plot = last_plot(), device = "bmp", path = NULL,
  scale = 1, width = 140, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)

#energy plot
posterior_cp <- as.array(datafit3)
np_cp <- nuts_params(datafit3)
color_scheme_set("blue")
mcmc_nuts_energy(np_cp)
ggsave("STAN/flights_model/figures/datafit3Figures/energyplot.bmp", plot = last_plot(), device = "bmp", path = NULL,
  scale = 1, width = 140, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)

#pairsplot (note, using just one theta to keep the plot readable)
theme_set(theme_classic(base_size = 7))
color_scheme_set("darkgray")

#mcmc_parcoord(posterior_cp, np = np_cp) #generates error: cannot allocate vector of size 889.3 Mb 
pairs_plot <- mcmc_pairs(posterior_cp, np = np_cp, pars = c("alpha","beta","lambda","kappa","theta[1]"), 
           off_diag_args = list(size = 0.75))

plots <- pairs_plot$bayesplots # list of 25 ggplot objects (diagonal plots are in slots 1,7,13,19,25)

#Note, to save the grid of the plots, set plots=bayesplot_grid(plots = plots)
ggsave("STAN/flights_model/figures/datafit3Figures/pairs_plot.jpeg", plot = bayesplot_grid(plots = plots), device = "jpeg", path = NULL,
  scale = 1, width = 140, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)

#pairsplot for Aki Vehtari with help from Jonah Gabry
posterior_cp <- posterior_cp%>%
  mutate(poserior_cp, lambda2 = qlogis)



pairs_plot2 <-
  mcmc_pairs(
    posterior_cp,
    np = np_cp,
    pars = c("alpha", "beta", "lambda", "kappa", "theta[1]"),
    transformations = list(
      "lambda" = "qlogis",  # or define a function called logit
      "kappa" = function(x) log(x - 0.1),
      "theta[1]" = "qlogis"
    ),
    off_diag_args = list(size = 0.75)
  )

plots2 <- pairs_plot2$bayesplots # list of 25 ggplot objects (diagonal plots are in slots 1,7,13,19,25)
plots2
#Note, to save the grid of the plots, set plots=bayesplot_grid(plots = plots)
ggsave("STAN/flights_model/figures/datafit3Figures/pairs_plot3.jpeg", plot = bayesplot_grid(plots = plots2), device = "jpeg", path = NULL,
  scale = 1, width = 140, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)

```
Result: consistent with model output, no divergences found. 

#5 Visualizing results
posteriors for known elodea-invaded lakes
```{r}
theme_set(theme_classic(base_size = 7))  #NOT WORKING HERE --> FIX!
library(ggplot2)
library(rstan)
FigPosteriors <- plot(datafit3, show_density = TRUE, pars=c("theta[321]","theta[322]","theta[324]","theta[376]","theta[518]","theta[300]","theta[309]","theta[376]","theta[347]","avg"), ci_level = 0.5, fill_color = "#99d8c9")
chvector <- character(length=9)
chvector <- c("Average", "Lake Hood","Daniels Lake","Stormy Lake","Alexander Lake","Sucker Lake","McKinley Lake","Martin Lake","Bering Lake")
FigPosteriors <- FigPosteriors + scale_y_continuous(labels=chvector[1:9],breaks=1:9) + labs(x="95% posterior for theta(j)")
FigPosteriors

ggsave("STAN/flights_model/figures/datafit3Figures/FigPosteriors.bmp", plot = last_plot(), device = "bmp", path = NULL,
  scale = 1, width = 140, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)

#Better publication version of the above, code not working correctly yet
posterior <- as.array(datafit3)
fig <- mcmc_areas(
  posterior, 
  pars = c("theta[321]","theta[322]","theta[324]","theta[376]","theta[518]","theta[300]","theta[309]","theta[376]","avg"),
  prob = 0.5, # 50% intervals
  prob_outer = 0.95, # 95%
  point_est = "median"
)

fig <- fig + scale_y_discrete(labels=chvector[1:9],breaks=1:9) + labs(x="95% posterior for theta(j)")
fig
```


Creating results table for seaplane bases
```{r}
FAAbases <- read.csv("Data/FAAbases.csv", stringsAsFactors = FALSE)
bases <- FAAbases[,2]
basesResults <- subset(results, UniqueID %in% bases)
```


Looking for divergences
```{r}
partition <- util$partition_div(datafit3)
div_samples <- partition[[1]]
nondiv_samples <- partition[[2]]

par(mfrow=c(1, 3))

plot(nondiv_samples$a, nondiv_samples$b,
     col=c_dark_trans, pch=16, cex=0.8,
     xlab="a", ylab="b")
points(div_samples$a, div_samples$b,
       col="green", pch=16, cex=0.8)

plot(nondiv_samples$a, log(nondiv_samples$sigma),
     col=c_dark_trans, pch=16, cex=0.8,
     xlab="a", ylab="log(sigma)")
points(div_samples$a, log(div_samples$sigma),
       col="green", pch=16, cex=0.8)

plot(nondiv_samples$b, log(nondiv_samples$sigma),
     col=c_dark_trans, pch=16, cex=0.8,
     xlab="b", ylab="log(sigma)")
points(div_samples$b, log(div_samples$sigma),
       col="green", pch=16, cex=0.8)
```
