---
title: "AIS introduction risk analysis"
author: ""
date: "December 30 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Hierarchical model using rstanarm for estimating destination specific probability distribution of introducing AIS to lake j  

##1 Set up and data import
Note, update url-links below for the following dataset archived at the NSF Arctic Data Center
Tobias Schwoerer. 2020. A survey of 2015 Alaska seaplane operations with predictions of transmission risk for aquatic invasive plants. Arctic Data Center.
```{r}
library(shinystan)
library(bayesplot)
library(ggplot2)
library(cowplot)
library(rstan)
library(dplyr)
library(tidyr)
library(foreign)
library(data.table)
library(tidyverse)
library(loo)


#importing data archived and update below links before running this Rmd file (for windows users)
flights <- read.csv(url("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3Af851caef-79de-4142-9b57-c2d7e68dd460",      method="libcurl"))
lakeList2  <-  read.csv(url("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3Aaa8e4e61-2d4a-481d-95c6-0901c269e7d8",method="libcurl"))

#Mac user try this to get the data
#flights <- read.csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3Af851caef-79de-4142-9b57-c2d7e68dd460")
#lakeList2  <-  read.csv("https://arcticdata.io/metacat/d1/mn/v2/object/urn%3Auuid%3Aaa8e4e61-2d4a-481d-95c6-0901c269e7d8")

```

##2 Data preparation
```{r}
#create ElodeaO and ElodeaD indicating whether origin or destination is known elodea source
starts <- filter(lakeList2, lakeList2$SeaplaneBase != "")
starts <- starts[,c("Elodea","SeaplaneBase")]
flights <- flights%>%
  left_join(starts,by=c("StartID"="SeaplaneBase"))
names(flights)[names(flights)=="Elodea"] <- "ElodeaO"
dest <- lakeList2[,c("Elodea","LakeID")]
flights <- flights%>%
  inner_join(dest,by=c("DestID"="LakeID"))
names(flights)[names(flights)=="Elodea"] <- "ElodeaD"

flights$Oeflights <- with(flights,ifelse(ElodeaO==0&ElodeaD==1|ElodeaO==1&ElodeaD==1,AnnualFl_1,0))
flights$Deflights <- with(flights,ifelse(ElodeaO==1&ElodeaD==0|ElodeaO==1&ElodeaD==1,AnnualFl_1,0))

#summarizing origin 
origin <- flights%>%
  group_by(StartID)%>%
  summarise(eflights=sum(Oeflights),
            flights=sum(AnnualFl_1))
ids <- lakeList2[,c("SeaplaneBase","LakeID")]
origin <- origin%>%  
  left_join(ids, by=c("StartID"="SeaplaneBase"))
#drop StartID column
origin <- origin%>%
  select(-StartID)

#summarizing destinations
destinations <- flights%>%
    group_by(DestID)%>%
    summarise(eflights=sum(Deflights),
            flights=sum(AnnualFl_1))
names(destinations)[names(destinations)=="DestID"] <- "LakeID"

#combining
data <- bind_rows(origin,destinations)%>%
  group_by(LakeID)%>%
    summarise(eflights=sum(eflights),
            flights=sum(flights))

J <- nrow(data)
n <- data$flights
y <- data$eflights
```

##3 Modeling
###3.1. Partial pooling model
Following: https://mc-stan.org/users/documentation/case-studies/pool-binary-trials-rstanarm.html#observed-vs.estimated-chance-of-success
```{r}
library(rstanarm)
Lake <- data$LakeID
lake_avgs <- y / n
tot_avg <- sum(y) / sum(n)
intro_avg <- function(x) print(format(round(x, digits = 3), nsmall = 3), quote = FALSE)
summary_stats <- function(posterior) {
  x <- invlogit(posterior)  # log-odds -> probabilities
  t(apply(x, 2, quantile, probs = c(0.1, 0.5, 0.9))) 
}
summary_mean <- function(posterior) {
  x <- invlogit(posterior)  # log-odds -> probabilities
  t(apply(x, 2, mean)) 
}

options(mc.cores = parallel::detectCores())
SEED <- 101
wi_prior <- normal(0, 1.4)  # this flat prior from King et al. (2009), also used weakly informative prior on log-odds normal(-1, 1), which is slightly informative on smaller values but compared to this flat prior was not as good in the model comparison below, see hist(plogis(rnorm(1000,-1,1)),100)

fit_partialpool <- 
  stan_glmer(cbind(y, n - y) ~ (1 | Lake), data = data,  family = binomial("logit"),
             prior_intercept = wi_prior, seed = SEED, adapt_delta = 0.99, iter=10000) #The left-hand side of the formula specifies the binomial outcome by providing the number of successes (hits) and failures (flights from non-sources) for each lake, and the right-hand side indicates that we want an intercept-only model.
```

###3.2. Complete pooling model and no pooling model
```{r}
fit_pool <- 
  stan_glm(cbind(y, n - y) ~ 1, data = data, family = binomial("logit"),
             prior_intercept = wi_prior, seed = SEED, adapt_delta = 0.99, iter=10000)

invlogit <- plogis  # function(x) 1/(1 + exp(-x))
summary_stats <- function(posterior) {
  x <- invlogit(posterior)  # log-odds -> probabilities
  t(apply(x, 2, quantile, probs = c(0.1, 0.5, 0.9))) 
}
summary_mean <- function(posterior) {
  x <- invlogit(posterior)  # log-odds -> probabilities
  t(apply(x, 2, mean)) 
}

pool <- summary_stats(as.matrix(fit_pool))  # as.matrix extracts the posterior draws

pool_mean <- summary_mean(as.matrix(fit_pool))  # as.matrix extracts the posterior draws

pool <- matrix(pool,  # replicate to give each player the same estimates
               nrow(data), ncol(pool), byrow = TRUE, 
               dimnames = list(data$LakeID, c("10%", "50%", "90%")))
intro_avg(pool)

#No pooling model
fit_nopool <- update(fit_pool, formula = . ~ 0 + Lake, prior = wi_prior)
nopool <- summary_stats(as.matrix(fit_nopool))
rownames(nopool) <- as.character(data$LakeID)
intro_avg(nopool)
```

##4 Model validation
Using cross-validation for model checking and comparison, approximating the expected log predictive density (the lower the better) in the third column of the output. Specifically, it is the leave-one-out (loo) approximation to the log predictive density. Even though this approximation is only asymptotically valid, as it likely underestimates the expected log predictive density, the relative ranking of the models is the same as if it would be correctly calculated. https://mc-stan.org/users/documentation/case-studies/pool-binary-trials-rstanarm.html#partial-pooling
```{r}
loo_compare(loo(fit_partialpool), loo(fit_pool), loo(fit_nopool))
```

##5 Plotting results and creating results table for archiving
```{r}
# shift each Lake's estimate by intercept (and then drop intercept)
shift_draws <- function(draws) {
  sweep(draws[, -1], MARGIN = 1, STATS = draws[, 1], FUN = "+")
}
alphas <- shift_draws(as.matrix(fit_partialpool))
partialpool <- summary_stats(alphas)
partialpool <- partialpool[-nrow(partialpool),]
rownames(partialpool) <- as.character(data$LakeID)
intro_avg(partialpool)

partialpool_mean <- summary_mean(alphas)[1,]
partialpool_mean = partialpool_mean[-length(partialpool_mean)]

#create dataframe and moving rownames into columns
partialpoolDF <- as.data.frame(partialpool)
partialpoolDF <- rownames_to_column(partialpoolDF, var="LakeID")

#save summary of fitted model as data.frame, remember this is in log-odds, so further down we add the above code which shifted results from logit to normal space. 
summary(fit_partialpool)
diag_partial <-as.data.frame(summary(fit_partialpool))
#moving rownames into column
diag_partial <- rownames_to_column(diag_partial, var="LakeID")
#extracting intercept, sigma, mean PPD, log-posterior to save for later
drop_rows <- diag_partial[(diag_partial$LakeID %in% c("(Intercept)","Sigma[Lake:(Intercept),(Intercept)]","mean_PPD","log-posterior")), ] 
#dropping the above
diag_partial <- diag_partial[ !(diag_partial$LakeID %in% c("(Intercept)","Sigma[Lake:(Intercept),(Intercept)]","mean_PPD","log-posterior")), ] #cleaning up LakeID by first extracting string that starts with Lk and ends with ]
diag_partial$LakeID <- gsub("b[(Intercept) Lake:","",diag_partial$LakeID, fixed=T)
diag_partial$LakeID <- substr(diag_partial$LakeID, 1, nchar(diag_partial$LakeID)-1)
#drop columns showing log-odds
diag_partial <- diag_partial %>%
  select(LakeID, mcse, n_eff, Rhat)

#joining dataframe to create results table incl. diagnostics, except sigma, mean PPD, log-posterior
partialpoolResults <- diag_partial%>%
  left_join(partialpoolDF, by="LakeID")

#diagnosing Gelman-Rubin statistic
partialpoolResults%>%
  summarise(RhatTotal =sum(Rhat>1.05), mcseTotal =sum(mcse>.1))

#writing results table as csv for online appendix to the paper and for archive in Arctic Data Center
resultsTable <- partialpoolResults%>%
  left_join(lakeList2, by="LakeID")
write.csv(resultsTable, file="/OnlineAppendix2.csv")
```

##6 Diagnostics for partial pooling model
```{r}
#Evaluating model convergence and diagnose model through STAN online portal
#launch_shinystan(fit_partialpool)
color_scheme_set("blue")
rhat_fig <- plot(fit_partialpool, "rhat")
rhat_fig
ess_fig <- plot(fit_partialpool, "ess")
ess_fig
ggsave("/FigA5.tiff", plot = last_plot(), device = "tiff", path = NULL,
  scale = 1, width = 90, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)

low_ess <- partialpoolResults%>%
  left_join(lakeList2, by="LakeID")%>%
  mutate(ess_N=n_eff/10000)%>%
  filter(ess_N<=0.1)

#number of seaplane bases with low ess
low_ess%>%
  filter(SeaplaneBase!="")%>%
  summarize(count=n())
```

##7 Graphical posterior predictive checks
see: https://mc-stan.org/users/documentation/case-studies/pool-binary-trials-rstanarm.html#graphical-posterior-predictive-checks
```{r}
#T stats plot
tstat_plots <- function(model, stats) {
  lapply(stats, function(stat) {
    graph <- pp_check(model, plotfun = "stat", stat = stat, 
                      seed = SEED) # optional arguments
    graph + xlab(stat) + theme(legend.position = "none")
  })
}
Tstats <- c("mean", "sd", "min","max")
#ppcs_nopool <- tstat_plots(fit_nopool, Tstats)
ppcs_partialpool <- tstat_plots(fit_partialpool, Tstats)

library(gridExtra)
grid.arrange(
  #arrangeGrob(grobs = ppcs_nopool, nrow = 1, left = "No Pooling"),
  arrangeGrob(grobs = ppcs_partialpool, nrow = 2)
)

ggsave("/FigA3.tiff", plot = last_plot(), device = "tiff", path = NULL,
  scale = 1, width = 90, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)

#Distributions of observed data and a random sample of replications
replicate <- pp_check(fit_partialpool, plotfun = "hist", nreps = 15, binwidth = 0.05) +
  xlab("AIS introduction rate") + ylab("Frequency") 
replicate
ggsave("/FigA4.tiff", plot = replicate, device = "tiff", path = NULL,
  scale = 1, width = 140, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)

#traceplot for convergence check after warm-up
color_scheme_set("mix-blue-red")
theme_set(theme_classic(base_size = 9))
traceplot <- mcmc_trace(fit_partialpool, pars = c("b[(Intercept) Lake:Lk_285505]",
                                                  "b[(Intercept) Lake:Lk_268]",
                                                  "b[(Intercept) Lake:Lk_243881]"),
           facet_args = list(nrow = 3)) 

levels(traceplot$data$parameter)[1] <- "Lake Hood"
levels(traceplot$data$parameter)[2] <- "Big Lake"
levels(traceplot$data$parameter)[3] <- "Martin Lake"
traceplot <- traceplot + xlab("iterations after warm up") +
                          scale_y_continuous(expression(paste( alpha[j])))
traceplot

#traceplot for the three lakes with the lowest effective sample size n_eff, also including divergent transitions
np_model <- nuts_params(fit_partialpool)  #checks and visualizes divergences in the traceplots
color_scheme_set("mix-blue-red")
theme_set(theme_classic(base_size = 9))
traceplot <- mcmc_trace(fit_partialpool, pars = c("b[(Intercept) Lake:Lk_285505]",
                                                  "b[(Intercept) Lake:Lk_315787]",
                                                  "b[(Intercept) Lake:Lk_36723]"),
                                        facet_args = list(nrow = 3),
                                        np=np_model) 
levels(traceplot$data$parameter)[1] <- "Lake Hood"
levels(traceplot$data$parameter)[2] <- "Big River Lakes"
levels(traceplot$data$parameter)[3] <- "Naknek Lake"
traceplot <- traceplot + xlab("iterations after warm up") +
                          scale_y_continuous(expression(paste( alpha[j])))
traceplot
ggsave("/FigA6.tiff", plot = traceplot, device = "tiff", path = NULL,
  scale = 1, width = 90, height = 140, units = "mm",
  dpi = 300, limitsize = TRUE)
```


#### Plotting predicted against the data, Posterior Medians and 80% Intervals
Below we also create a reference table associated with all the AIS invaded lakes
```{r}
library(ggplot2)
models <- c("no pooling","partial pooling")  #"complete pooling"
estimates <- rbind(nopool,partialpool)  #pool, nopool, 
colnames(estimates) <- c("lb", "median", "ub")
plotdata <- data.frame(estimates, 
                       observed = rep(lake_avgs, times = length(models)), 
                       model = rep(models, each = J))
plotdata$LakeID <- rownames(plotdata)
#cleaning up LakeID due to rbind command above
plotdata$LakeID <- with(plotdata,ifelse(model=="partial pooling",substr(plotdata$LakeID, 1, nchar(plotdata$LakeID)-2),plotdata$LakeID))
plotdata <- plotdata%>%
  left_join(elodeaLakes,by="LakeID")
highlight_plot <- plotdata %>% 
             filter(Elodea==1)
              
#comparison plot
compareModels <- ggplot(plotdata, aes(x = observed, y = median, ymin = lb, ymax = ub)) +
  geom_abline(intercept = 0, slope = 1, color = "skyblue") + 
  geom_linerange(color = "gray60", size = 0.75) + 
  geom_point(size = 2.5, shape = 21, fill = "gray30", color = "white", stroke = 0.2) + 
  facet_grid(. ~ model) +
  coord_fixed() +
  scale_x_continuous(breaks = c(0.25,0.50,0.75,1.0)) +
  labs(x = "Observed y / n", y = expression(paste("Predicted, ",theta["j"])))
compareModels
ggsave("/FigA2.tiff", plot = compareModels, device = "tiff", path = NULL,
  scale = 1, width = 140, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)
```


#### Plotting posterior distributions
creating reference tables for later analysis and plotting
```{r}
elodeaLakes <- lakeList2%>%
  filter(Elodea==1 & LakeName !="Tanana River")%>%
  select(LakeID,Elodea,LakeName)%>%
  mutate(modelID=paste("b[(Intercept) Lake:",LakeID,"]", sep=''))%>%
  mutate(LakeName=replace(LakeName, LakeName=="Lake Hood/Lake Spenard", "Lake Hood"))

spBases <- unique(lakeList2$SeaplaneBase[grep("^S", lakeList2$SeaplaneBase)])
SpBases_List <- subset(lakeList2, SeaplaneBase %in% spBases)
```

1. posterior plots by region
facet_grid did not show the data nicely, so created individual plots for each region
More code but nicer plot compared to using facet-grid
```{r}
allData <- partialpoolDF
colnames(allData)[2] <- "llimit"
colnames(allData)[3] <- "median"
colnames(allData)[4] <- "ulimit"

# bring in mean
allData$mean = partialpool_mean
# also bring in raw data -- flight #s
allData = dplyr::left_join(allData, data)

allData <- allData%>%
  left_join(lakeList2,by=c("LakeID"))%>%
  mutate(LakeNameID=paste(LakeName," ",LakeID))
theme_set(theme_classic(base_size = 9))

#Bristol Bay
BB <- allData%>%
  filter(Region=="Bristol Bay")
BBplot <- BB %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Bristol Bay") +xlim(0, 1)
BBplot

#Cook Inlet
CI <- allData%>%
  filter(Region=="Cook Inlet")
CIplot <- CI %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Cook Inlet") +xlim(0, 1)
CIplot

#Gulf
Gulf <- allData%>%
  filter(Region=="Gulf")
Gulfplot <- Gulf %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Gulf") +xlim(0, 1)
Gulfplot

#Kuskokwim
Kusko <- allData%>%
  filter(Region=="Kuskokwim")
Kuskoplot <- Kusko %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Kuskokwim") + xlim(0, 1)
Kuskoplot

#North Slope
NS <- allData%>%
  filter(Region=="North Slope")
NSplot <- NS %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("North Slope") +xlim(0, 1)
NSplot

#Yukon
YK <- allData%>%
  filter(Region=="Yukon")
YKplot <- YK %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2)+
  ylab("waterbodies") + xlab("posterior predictive distribution") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Yukon") +xlim(0, 1)
YKplot

#Knik Arm
KA <- allData%>%
  filter(Region=="Knik Arm")
KAplot <- KA %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Knik Arm") +xlim(0, 1)
KAplot

#Kodiak
K <- allData%>%
  filter(Region=="Kodiak")
Kplot <- K %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,median)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameID, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Kodiak") +xlim(0, 1)
Kplot

library(cowplot)
theme_set(theme_cowplot(font_size=7))
RegionPlot <-plot_grid(CIplot,Gulfplot,KAplot,Kplot,Kuskoplot,BBplot,YKplot,NSplot,  ncol = 2, align = "h", axis = "bt", rel_widths = c(1, 1) )
RegionPlot
ggsave("/Fig3.tiff", plot = RegionPlot, device = "tiff",
  path = NULL,
  scale = 1, width = 140, height = 200, units = "mm",
  dpi = 300, limitsize = TRUE)
```

```{r}
# Eric added this plot to swap in the  mean 
#Bristol Bay
BB <- allData%>%
  filter(Region=="Bristol Bay")
BBplot <- BB %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Bristol Bay")
BBplot

#Cook Inlet
CI <- allData%>%
  filter(Region=="Cook Inlet")
CIplot <- CI %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Cook Inlet")
CIplot

#Gulf
Gulf <- allData%>%
  filter(Region=="Gulf")
Gulfplot <- Gulf %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Gulf")
Gulfplot

#Kuskokwim
Kusko <- allData%>%
  filter(Region=="Kuskokwim")
Kuskoplot <- Kusko %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Kuskokwim")
Kuskoplot

#North Slope
NS <- allData%>%
  filter(Region=="North Slope")
NSplot <- NS %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("North Slope")
NSplot

#Yukon
YK <- allData%>%
  filter(Region=="Yukon")
YKplot <- YK %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("waterbodies") + xlab("posterior predictive distribution") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Yukon")
YKplot

#Knik Arm
KA <- allData%>%
  filter(Region=="Knik Arm")
KAplot <- KA %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Knik Arm")
KAplot

#Kodiak
K <- allData%>%
  filter(Region=="Kodiak")
Kplot <- K %>%
  mutate(LakeNameID=fct_reorder(LakeNameID,mean*flights)) %>%  
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameID, xmin = llimit*flights, xmax = ulimit*flights), shape=19,fatten = 4, size = .2)+
  ylab("") + xlab("") +
  theme(axis.text.y = element_blank(),axis.ticks = element_blank())+
  ggtitle("Kodiak")
Kplot

library(cowplot)
theme_set(theme_cowplot(font_size=7))
RegionPlot <-plot_grid(CIplot,Gulfplot,KAplot,Kplot,Kuskoplot,BBplot,YKplot,NSplot,  ncol = 2, align = "h", axis = "bt", rel_widths = c(1, 1) )
RegionPlot
ggsave("/Fig3_b.tiff", plot = RegionPlot, device = "tiff",
  path = NULL,
  scale = 1, width = 140, height = 200, units = "mm",
  dpi = 300, limitsize = TRUE)
```


Note, the invasion scenario for the Fairbanks Floatpond used the same code as above after in lakeList2 changing the Elodea variable for Fairbanks Floatpond from 0 to 1 for the scenario run, then saving the Yukon and North Slope as Fig4.   

2. Elodea invaded lakes
```{r}
ElodeaData <- partialpoolDF%>%
  left_join(lakeList2,by=c("LakeID"))%>%
  filter(Elodea==1 & LakeName!="Tanana River")
colnames(ElodeaData)[2] <- "llimit"
colnames(ElodeaData)[3] <- "median"
colnames(ElodeaData)[4] <- "ulimit"

EPlot<- ElodeaData %>%
  mutate(LakeName=fct_reorder(LakeName,median)) %>%  #fct_reorder guarantees correct ordering of the labels with data in the plot
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeName, xmin = llimit, xmax = ulimit))+
  #geom_point(data = pick(~Elodea == 1), colour = "red")+
  ylab("") + xlab("posterior predictive distribution") 
EPlot
ggsave("/EPlot.tiff", plot = EPlot, device = "tiff", path = NULL,
  scale = 1, width = 90, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)
```

```{r}
ppoolDF = partialpoolDF
ppoolDF$mean = partialpool_mean
ppoolDF = dplyr::left_join(ppoolDF, data)

ElodeaData <- ppoolDF%>%
  left_join(lakeList2,by=c("LakeID"))%>%
  filter(Elodea==1 & LakeName!="Tanana River")
names(ElodeaData)[2] <- "llimit"
names(ElodeaData)[3] <- "median"
names(ElodeaData)[4] <- "ulimit"

EPlot<- ElodeaData %>%
  mutate(LakeName=fct_reorder(LakeName,mean*flights)) %>%  #fct_reorder guarantees correct ordering of the labels with data in the plot
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeName, xmin = llimit*flights, xmax = ulimit*flights))+
  #geom_point(data = pick(~Elodea == 1), colour = "red")+
  ylab("") + xlab("posterior predictive distribution") 
EPlot
ggsave("/EPlot_b.tiff", plot = EPlot, device = "tiff", path = NULL,
  scale = 1, width = 90, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)
```

3. Seaplane bases
creating subset for seaplane bases, showing 80% posterior intervals
```{r}
spData <- partialpoolDF%>%
  filter(LakeID%in%SpBases_List$LakeID)%>%
  left_join(lakeList2,by=c("LakeID"))%>%
  mutate(LakeNameCity=paste(LakeName," ",SPBaseCity))
colnames(spData)[2] <- "llimit"
colnames(spData)[3] <- "median"
colnames(spData)[4] <- "ulimit"

spPlot<- spData %>%
  mutate(LakeNameCity=fct_reorder(LakeNameCity,median)) %>%
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameCity, xmin = llimit, xmax = ulimit),shape=19,fatten = 4, size = .2,color=ifelse(spData$Elodea==1, 'red', 'black'))+
  #geom_point(data = pick(~Elodea == 1), colour = "red")+
  ylab("") + xlab("posterior predictive distribution") 
spPlot
ggsave("/FigA7.tiff", plot = spPlot, device = "tiff", path = NULL,
  scale = 1, width = 140, height = 200, units = "mm",
  dpi = 300, limitsize = TRUE)
```

```{r}
ppoolDF = partialpoolDF
ppoolDF$mean = partialpool_mean
ppoolDF = dplyr::left_join(ppoolDF, data)

spData <- ppoolDF%>%
  filter(LakeID%in%SpBases_List$LakeID)%>%
  left_join(lakeList2,by=c("LakeID"))%>%
  mutate(LakeNameCity=paste(LakeName," ",SPBaseCity))
colnames(spData)[2] <- "llimit"
colnames(spData)[3] <- "median"
colnames(spData)[4] <- "ulimit"

spPlot<- spData %>%
  mutate(LakeNameCity=fct_reorder(LakeNameCity,mean*flights)) %>%
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameCity, xmin = llimit*flights, xmax = ulimit*flights),shape=19,fatten = 4, size = .2,color=ifelse(spData$Elodea==1, 'red', 'black'))+
  #geom_point(data = pick(~Elodea == 1), colour = "red")+
  ylab("") + xlab("posterior predictive distribution") 
spPlot
ggsave("/FigA7_b.tiff", plot = spPlot, device = "tiff", path = NULL,
  scale = 1, width = 140, height = 200, units = "mm",
  dpi = 300, limitsize = TRUE)
```

4. combined figure with elodea infestations and top 10 seaplane bases
```{r}
#seaplane bases
spDataTop10 <- partialpoolDF%>%
  filter(LakeID%in%SpBases_List$LakeID)%>%
  left_join(lakeList2,by=c("LakeID"))%>%
  mutate(LakeNameRegion=paste(LakeName," ",Region))
colnames(spDataTop10)[2] <- "llimit"
colnames(spDataTop10)[3] <- "median"
colnames(spDataTop10)[4] <- "ulimit"
spDataTop10 <- top_n(spDataTop10,10,median)
  
#elodea infestations
ElodeaData2 <- ElodeaData%>%
  mutate(LakeNameRegion=paste(LakeName," ",Region))
#combining the two above and creating a seaplane base variable
prediPlotData <- union(spDataTop10,ElodeaData2)
prediPlotData$Seaplane_Base <- ifelse(prediPlotData$SeaplaneBase!='',"yes","no")
prediPlotData$Elodea_present <- ifelse(prediPlotData$Elodea==1,"yes","no")

#second try with legend
prediPlot <- prediPlotData %>%
  mutate(LakeNameRegion=fct_reorder(LakeNameRegion,median)) %>%
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = median, y = LakeNameRegion, xmin = llimit, xmax = ulimit, shape=Seaplane_Base,color=Elodea_present ))+
  ylab("") + xlab("posterior predictive distribution")+
    scale_color_discrete(name = "Elodea present") +
  scale_shape_discrete(name = "Seaplane Base") +
  theme(legend.position = c(.2,.7))+
  scale_color_viridis(discrete=TRUE)
prediPlot
ggsave("/Fig2.tiff", plot = prediPlot, device = "tiff", path = NULL,
  scale = 1, width = 140, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)
```

```{r}
ppoolDF = partialpoolDF
ppoolDF$mean = partialpool_mean
ppoolDF = dplyr::left_join(ppoolDF, data)

#seaplane bases
spDataTop10 <- ppoolDF%>%
  filter(LakeID%in%SpBases_List$LakeID)%>%
  left_join(lakeList2,by=c("LakeID"))%>%
  mutate(LakeNameRegion=paste(LakeName," ",Region))
colnames(spDataTop10)[2] <- "llimit"
colnames(spDataTop10)[3] <- "median"
colnames(spDataTop10)[4] <- "ulimit"
spDataTop10 <- top_n(spDataTop10,10,mean*flights)
  
#elodea infestations
ElodeaData2 <- ElodeaData%>%
  mutate(LakeNameRegion=paste(LakeName," ",Region))
#combining the two above and creating a seaplane base variable
prediPlotData <- union(spDataTop10,ElodeaData2)
prediPlotData$Seaplane_Base <- ifelse(prediPlotData$SeaplaneBase!='',"yes","no")
prediPlotData$Elodea_present <- ifelse(prediPlotData$Elodea==1,"yes","no")

#second try with legend
prediPlot <- prediPlotData %>%
  mutate(LakeNameRegion=fct_reorder(LakeNameRegion,mean*flights)) %>%
  ggplot()+
  theme_set(theme_classic(base_size = 9))+
  geom_pointrange(aes(x = mean*flights, y = LakeNameRegion, xmin = llimit*flights, xmax = ulimit*flights, shape=Seaplane_Base,color=Elodea_present ))+
  ylab("") + xlab("posterior predictive distribution")+
    scale_color_discrete(name = "Elodea present") +
  scale_shape_discrete(name = "Seaplane Base") +
  theme(legend.position = c(.2,.7))+
  scale_color_viridis(discrete=TRUE)
prediPlot
ggsave("/Fig2_b.tiff", plot = prediPlot, device = "tiff", path = NULL,
  scale = 1, width = 140, height = 90, units = "mm",
  dpi = 300, limitsize = TRUE)
```


#### Regional statistics
```{r}
#Top Seaplane bases by region, first matching Destination ID with LakeID to get the destination region
topSPB <- flights%>%
  left_join(lakeList2,by=c("StartID"="SeaplaneBase"))%>%
  select(LakeName,DestID,AnnualFl_1,Region)
colnames(topSPB)[4] <- "StRegion"
  topSPB <- topSPB%>%
  left_join(select(lakeList2,LakeID,Region),by=c("DestID"="LakeID"))

#Ranking of seaplane bases according to flight trip frequency per region and total 
topSPB2 <- topSPB%>%
  group_by(LakeName, StRegion,Region)%>%
  summarize(total = sum(AnnualFl_1))%>%
  spread(Region,total)
write.csv(topSPB,"/OnlineAppendix3.csv")  
```

